---
title: "Assignment 3 - Part 2 - Diagnosing Schizophrenia from Voice"
author: "Riccardo Fusaroli"
date: "October 17, 2017"
output: 
  md_document:
    variant: markdown_github 
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(pacman, tidyverse, tidymodels, groupdata2, lmerTest, lme4, kernlab, ggplot2)
allData <- read.csv('full_data.csv')
```

## Assignment 3 - Part 2 - Diagnosing schizophrenia from voice

In the previous part of the assignment you generated a bunch of "features", that is, of quantitative descriptors of voice in schizophrenia. We then looked at whether we could replicate results from the previous literature.
We now want to know whether we can automatically diagnose schizophrenia from voice only, that is, relying on the set of features you produced last time, we will try to produce an automated classifier.
Again, remember that the dataset containst 7 studies and 3 languages. Feel free to only include Danish (Study 1-4) if you feel that adds too much complexity.

Issues to be discussed your report:
- Should you run the analysis on all languages/studies at the same time? 
- Choose your best acoustic feature from part 1. How well can you diagnose schizophrenia just using it?
- Identify the best combination of acoustic features to diagnose schizophrenia using logistic regression.
- Discuss the "classification" process: which methods are you using? Which confounds should you be aware of? What are the strength and limitation of the analysis?
- Bonus question: Logistic regression is only one of many classification algorithms. Try using others and compare performance. Some examples: Discriminant Function, Random Forest, Support Vector Machine, etc. The package caret provides them. 
- Bonus Bonus question: It is possible combine the output of multiple  classification models to improve classification accuracy. For inspiration see,
https://machinelearningmastery.com/machine-learning-ensembles-with-r/
 The interested reader might also want to look up 'The BigChaos Solution to the Netflix Grand Prize'

## Learning objectives
- Learn the basics of classification in a machine learning framework
- Design, fit and report logistic regressions
- Apply feature selection techniques

### Let's start

We first want to build a logistic regression to see whether you can diagnose schizophrenia from your best acoustic feature. Let's use the full dataset and calculate the different performance measures (accuracy, sensitivity, specificity, PPV, NPV, ROC curve). You need to think carefully as to how we should (or not) use study and subject ID.

Then cross-validate the logistic regression and re-calculate performance on the testing folds. N.B. The cross-validation functions you already have should be tweaked: you need to calculate these new performance measures. Alternatively, the groupdata2 and cvms package created by Ludvig are an easy solution. 

N.B. the predict() function generates log odds (the full scale between minus and plus infinity). Log odds > 0 indicates a choice of 1, below a choice of 0.
N.N.B. you need to decide whether calculate performance on each single test fold or save all the prediction for test folds in one datase, so to calculate overall performance.
N.N.N.B. Now you have two levels of structure: subject and study. Should this impact your cross-validation?
N.N.N.N.B. A more advanced solution could rely on the tidymodels set of packages (warning: Time-consuming to learn as the documentation is sparse, but totally worth it)



```{r}
# Making a dataframe grouped by participant 
# Grouping each variable by participant
grouped_iqr <- aggregate(x = allData$IQR, by = list(allData$uID), FUN = mean, na.rm=T) %>% dplyr::rename(IQR=x)
grouped_sd <- aggregate(x = allData$sd, by = list(allData$uID), FUN = mean, na.rm=T) %>% dplyr::rename(sd=x)
grouped_pauseDuration <- aggregate(x = allData$pauseDuration, by = list(allData$uID), FUN = mean, na.rm=T) %>% dplyr::rename(pauseDuration=x)
grouped_propSpokenTime <- aggregate(x = allData$propSpokenTime, by = list(allData$uID), FUN = mean, na.rm=T) %>% dplyr::rename(propSpokenTime=x)
grouped_speechrate <- aggregate(x = allData$speechrate..nsyll.dur., by = list(allData$uID), FUN = mean, na.rm=T) %>% dplyr::rename(speechrate=x)
grouped_diagnosis <- aggregate(x = as.numeric(allData$diagnosis), by = list(allData$uID), FUN = mean, na.rm=T) %>% dplyr::rename(diagnosis=x)
grouped_study <- aggregate(x = as.numeric(allData$study), by = list(allData$uID), FUN = mean, na.rm=T) %>% dplyr::rename(study=x)


# Merged grouped data frames
merged_grouped <- merge(grouped_iqr, grouped_pauseDuration, by="Group.1")
merged_grouped <- merge(merged_grouped, grouped_propSpokenTime, by="Group.1")
merged_grouped <- merge(merged_grouped, grouped_speechrate, by="Group.1")
merged_grouped <- merge(merged_grouped, grouped_sd, by="Group.1")
merged_grouped <- merge(merged_grouped, grouped_study, by="Group.1")
merged_grouped <- merge(merged_grouped, grouped_diagnosis, by="Group.1")

# Changing Group.1 collumn name to uID
groupedData <- merged_grouped %>% dplyr::rename(uID=Group.1)

# Saving grouped data as csv file
write_csv(merged_grouped, "groupedData.csv")

# loading grouped data
groupedData$diagnosis <- as.factor(groupedData$diagnosis)
```



# Defining models with grouped data (1 data point pr. participant instead of approx. 10)
```{r}
# creating dataframes including relevant variables
groupedDataIQR <- groupedData %>% 
  select(uID, diagnosis, study, IQR) %>% drop_na() # Pitch variablity model

groupedDataPause <- groupedData %>% 
  select(uID, diagnosis, study, pauseDuration) %>% drop_na() # Pause duration model

groupedDataSpeechrate <- groupedData %>% 
  select(uID, diagnosis, study, speechrate) %>% drop_na() # Speech rate model

groupedDataSpokentime <- groupedData %>% 
  select(uID, diagnosis, study, propSpokenTime) %>% drop_na() # Proportional of spoken time model

groupedDataSD <-  groupedData %>% 
  select(uID, diagnosis, study, sd) %>% drop_na() # Standard deviation model

groupedData1 <- groupedData %>% 
  select(uID, diagnosis, study, IQR, pauseDuration, 
         speechrate, propSpokenTime) %>% drop_na() 

groupedData2 <- groupedData %>% 
  select(uID, diagnosis, study, IQR, pauseDuration, 
         speechrate, sd) %>% drop_na() 

groupedData3 <- groupedData %>% 
  select(uID, diagnosis, study, IQR, pauseDuration, 
         propSpokenTime, sd) %>% drop_na() 

groupedData4 <- groupedData %>% 
  select(uID, diagnosis, study, IQR, 
         speechrate, propSpokenTime, sd) %>% drop_na() 

groupedData5 <-  groupedData %>% 
  select(uID, diagnosis, study, pauseDuration, 
         speechrate, propSpokenTime, sd) %>% drop_na() 

groupedDataCombined <- groupedData %>% 
  select(uID, diagnosis, study, IQR, pauseDuration, 
         speechrate, propSpokenTime, sd) %>% drop_na() # The combined model

```



# Running models & cross-validation
```{r}
# partitioning the data using groupdata2 by diagnosis and study as cat_cols
df_list <- partition(groupedDataSD, p = 0.2, cat_col = c("diagnosis",'study'), list_out = T)

 # defining the test-set and removing ID and study column
df_test <- df_list[[1]]
df_test <- df_test %>% 
  select(-uID, -study)

# defining the train-set and removing ID-column
df_train <- df_list[[2]]
df_train <- df_train %>% 
  select(-uID, -study)



# Defining the recipe for train-data
# We've removed NAs so no need to check for missing values
rec <- df_train %>% recipe(diagnosis ~ .) %>% # defines outcome of pre-processing
  step_center(all_numeric()) %>% # centering all numeric values
  step_scale(all_numeric()) %>% # scaling all numeric values
  step_corr(all_numeric()) %>% # corr testing all predictors
  prep(training = df_train) # defining the train-set

# extracting 'df_train' from rec_train (the recipe)
train_baked <- juice(rec)
test_baked <- rec %>% bake(df_test)


# defining model as logistic regression
log_fit <- 
  logistic_reg() %>% 
  set_mode("classification") %>% 
  set_engine("glm") %>% 
  fit(diagnosis ~ ., data = train_baked)

# defining model as a support-vector-machine
svm_fit <- 
  svm_rbf() %>% 
  set_mode("classification") %>% 
  set_engine("kernlab") %>%
  fit(diagnosis ~ ., data = train_baked)



# investigating both logistic regression and the SVM-models
# get multiple at once
test_results <- 
  test_baked %>% 
  select(diagnosis) %>% 
  mutate(
    log_class = predict(log_fit, new_data = test_baked) %>% 
      pull(.pred_class),
    log_prob  = predict(log_fit, new_data = test_baked, type = "prob") %>% 
      pull(.pred_1),
    svm_class = predict(svm_fit, new_data = test_baked) %>% 
      pull(.pred_class),
    svm_prob  = predict(svm_fit, new_data = test_baked, type = "prob") %>% 
      pull(.pred_1)
  )


# Investigating the metrics
metrics(test_results, truth = diagnosis, estimate = log_class) %>% 
  knitr::kable()

# extracting metrics of log_class
test_results %>%
  select(diagnosis, log_class, log_prob) %>%
  knitr::kable()


# # Plotting area-under-the-curve (ROC-curve)
test_results %>%
  roc_curve(truth = diagnosis, log_prob) %>%
  autoplot()

# Cross validation of grouped data
cv_folds <- vfold_cv(df_train, v = 5, repeats = 2, strata = diagnosis, group = uID)

#prepare data set and fetch train data
cv_folds <- cv_folds %>% 
  mutate(recipes = splits %>%
           # prepper is a wrapper for `prep()` which handles `split` objects
           map(prepper, recipe = rec),
         train_data = splits %>% map(training))

# train model of each fold
  # create a non-fitted model
log_fit <- 
  logistic_reg() %>%
  set_mode("classification") %>% 
  set_engine("glm")

cv_folds <- cv_folds %>%  mutate(
  log_fits = pmap(list(recipes, train_data), #input 
                            ~ fit(log_fit, formula(.x), data = bake(object = .x, new_data = .y)) # function to apply
                 ))

predict_log <- function(split, rec, model) {
  # IN
    # split: a split data
    # rec: recipe to prepare the data
    # 
  # OUT
    # a tibble of the actual and predicted results
  baked_test <- bake(rec, testing(split))
  tibble(
    actual = baked_test$diagnosis,
    predicted = predict(model, new_data = baked_test) %>% pull(.pred_class),
    prop_sui =  predict(model, new_data = baked_test, type = "prob") %>% pull(.pred_1),
    prop_non_sui =  predict(model, new_data = baked_test, type = "prob") %>% pull(`.pred_0`)
  ) 
}

# apply our function to each split, which their respective recipes and models (in this case log fits) and save it to a new col
cv_folds <- cv_folds %>% 
  mutate(pred = pmap(list(splits, recipes, log_fits) , predict_log))


eval <- cv_folds %>% 
  mutate(
    metrics = pmap(list(pred), ~ metrics(., truth = actual, estimate = predicted, prop_sui))) %>% 
  select(id, id2, metrics) %>% 
  unnest(metrics)

# inspect performance metrics
 eval %>% 
  select(repeat_n = id, fold_n=id2,metric = .metric, estimate = .estimate) %>% 
  spread(metric, estimate) %>% 
  head() %>% 
  knitr::kable()
```




# Making a combined prediction from multiple models
```{r}
# List of multiple models
dataList = list(groupedData1, groupedData2, groupedData3,
            groupedData4, groupedData5, groupedDataCombined)

# empty data frame
allPredictions <- data.frame()


# Creating loop to run through list of models
for (df in dataList){ #Partitioning data
  df_list <- partition(df, p = 0.2, cat_col = c("diagnosis", "study"), list_out = T)
  
  # define test-set
  df_test <- df_list[[1]]
  df_test <- df_test %>% 
  select(-uID, -study)

  # define train-set
  df_train <- df_list[[2]]
  df_train <- df_train %>% 
  select(-uID, -study)
  
  # Defining the recipe for train-data
  # We've removed NAs so no need to check for missing values
  rec <- df_train %>% recipe(diagnosis ~ .) %>% # defines outcome of pre-processing
    step_center(all_numeric()) %>% # centering all numeric values
    step_scale(all_numeric()) %>% # scaling all numeric values
    step_corr(all_numeric()) %>% 
    prep(training = df_train) # defining the train-set

  # extracting 'df_train' from rec_train (the recipe)
  train_baked <- juice(rec)
  test_baked <- rec %>% bake(df_test)


  # defining model as logistic regression
  log_fit <- 
    logistic_reg() %>% 
    set_mode("classification") %>% 
    set_engine("glm") %>% 
    fit(diagnosis ~ ., data = train_baked)
  
  # defining model as a support-vector-machine
  svm_fit <- 
    svm_rbf() %>% 
    set_mode("classification") %>% 
    set_engine("kernlab") %>%
    fit(diagnosis ~ ., data = train_baked)
  
  # getting test-results
  test_results <- test_baked %>% 
    select(diagnosis) %>% 
    mutate(
      log_class = predict(log_fit, new_data = test_baked) %>% 
        pull(.pred_class),
      log_prob  = predict(log_fit, new_data = test_baked, type = "prob") %>% 
        pull(.pred_1),
      svm_class = predict(svm_fit, new_data = test_baked) %>% 
        pull(.pred_class),
      svm_prob  = predict(svm_fit, new_data = test_baked, type = "prob") %>% 
        pull(.pred_1)
    )
  
    #  Investigating the metrics
  metrics(test_results, truth = diagnosis, estimate = log_class) %>% 
    knitr::kable()
  
  # extracting metrics of log_class
  test_results <- test_results %>%
    select(diagnosis, log_class, log_prob) %>% 
    mutate(log_prob = log_prob-0.5,
           diagnosis = as.numeric(diagnosis)-1,
           log_class = as.numeric(log_class)-1) 
  
  if (NROW(allPredictions) < 1)  { #saving test results into a dataframe
        allPredictions <- test_results
  } else {
    allPredictions <- cbind(allPredictions, test_results)
  }
}

# Function for extracting max deviation from 0.5
absmax <- function(x) {x[which.max ( abs(x) )]}

# creating empty list
ulala <- list()
  
# add max deviation value of probabilty of prediction between the five models from each participant into a list
for (i in 1:44){
  ulala[i] <- absmax(allPredictions[i, c(3, 6, 9, 12, 15, 18)])
}

# transposing list into a vector
ulala <- unlist(transpose(ulala))

# Make new data fram with absolute values of deviations from 0.5
absAllPredictions <- allPredictions
absAllPredictions[, c(3, 6, 9, 12, 15, 18)] <- abs(allPredictions[, c(3, 6, 9, 12, 15, 18)])

# Summing probability outputs for each prediction from each model for each participant
sumLogProb = absAllPredictions[,3]+absAllPredictions[,6]+absAllPredictions[,9]+absAllPredictions[,12]+absAllPredictions[,15]+absAllPredictions[,18]

# Calculating weighted sum of probabilities
final <- absAllPredictions[,2]*(absAllPredictions[,3]/sumLogProb)+absAllPredictions[,5]*(absAllPredictions[,6]/sumLogProb)+absAllPredictions[,8]*(absAllPredictions[,9]/sumLogProb)+absAllPredictions[,11]*(absAllPredictions[,12]/sumLogProb)+absAllPredictions[,14]*(absAllPredictions[,15]/sumLogProb)+absAllPredictions[,17]*(absAllPredictions[,18]/sumLogProb)

# Adding predictions from all 6 models together
simplePred <- absAllPredictions[,2]+absAllPredictions[,5]+absAllPredictions[,8]+absAllPredictions[,11]+absAllPredictions[,14]+absAllPredictions[,17]

# Adding 'weighted sum' probs and actual predictions to a data frame
predData <- data.frame(predictions=final, actual=allPredictions[,1])
# Turning weighted sum probabilities into diagnosis predictions
predData$predDiagnosis <- ifelse(predData$predictions < 0.5, "0", "1")
predData$predDiagnosis <- as.factor(predData$predDiagnosis)
predData$actual <- as.factor(predData$actual)

# Making confusion matrix
caret::confusionMatrix(predData$predDiagnosis, predData$actual, positive="0")

# Same for max prob method
predData2 <- data.frame(predictions=ulala, actual=allPredictions[,1])
predData2$predDiagnosis <- ifelse(predData2$predictions < 0, "0", "1")
predData2$predDiagnosis <- as.factor(predData2$predDiagnosis)
predData2$actual <- as.factor(predData2$actual)

caret::confusionMatrix(predData2$predDiagnosis, predData2$actual, positive="0")

#Same for un-weighted sum method
predData3 <- data.frame(predictions=simplePred, actual=allPredictions[,1])
predData3$predDiagnosis <- ifelse(predData3$predictions < 3 , "0", "1")
predData3$predDiagnosis <- as.factor(predData3$predDiagnosis)
predData3$actual <- as.factor(predData3$actual)

caret::confusionMatrix(predData3$predDiagnosis, predData3$actual, positive="0")

# Combining all three methods
prediction_df <- data.frame(predWeight=as.numeric(predData$predDiagnosis)-1, predMax=as.numeric(predData2$predDiagnosis)-1, predSum=as.numeric(predData3$predDiagnosis)-1, actual=allPredictions[,1])
prediction_df$predDiagnosis <- ifelse(prediction_df$predWeight+prediction_df$predMax+prediction_df$predSum < 2, "0", "1")
prediction_df$predDiagnosis <- as.factor(prediction_df$predDiagnosis)
prediction_df$actual <- as.factor(prediction_df$actual)

caret::confusionMatrix(prediction_df$predDiagnosis, prediction_df$actual, positive="0")

pacman::p_load(precrec)

# Plotting roc-curve for weighted sum method 
precrec_obj <- evalmod(scores = as.numeric(predData$predDiagnosis)-1, labels = as.numeric(predData$actual)-1)
autoplot(precrec_obj)

citation("precrec")
```

```{r}
# Attempting to stack multiple methods
p_load(caretEnsemble, caret, randomForest)
train_ensemble <- groupedData %>% select(-uID, -study)
train_ensemble <- na.omit(train_ensemble)
train_ensemble$diagnosis <- as.factor(train_ensemble$diagnosis)
str(train_ensemble)
levels(train_ensemble$diagnosis) <- c('control', 'schizo')

control <- trainControl(method="repeatedcv", number=5, repeats=2, savePredictions=TRUE, classProbs=TRUE)
algorithmList <- c('lda','svmRadial','rpart', 'glm', 'knn')
modls <- caretList(diagnosis~. , data=train_ensemble, trControl=control, methodList=algorithmList)
results <- resamples(modls)
summary(results)
dotplot(results)

# correlation between results
modelCor(results)
splom(results)

# stack using glm
stackControl <- trainControl(method="repeatedcv", number=5, repeats=2, savePredictions=TRUE, classProbs=TRUE)
stack.glm <- caretStack(modls, method="glm", metric="Accuracy", trControl=stackControl)
print(stack.glm)

# stack using random forest
stack.rf <- caretStack(modls, method="rf", metric="Accuracy", trControl=stackControl)
print(stack.rf)
```

